TacticToe: Learning the Game of Tactics

### 2018-01-30 tactictoe_jv.tex
fixing conflicting notations in Abstraction

\paragraph{Abstraction of Tactic Arguments}
The first step is to abstract arguments of tactics by replacing them by a 
place holder creating an tactic with an unspecified argument.
Suppose we have a recorded tactic $T$. Since $T$ is a \sml code tree, we can 
try to abstract any of the 
\sml subterms. Let $u$ be a \sml subterm of $T$, and $h$ a placeholder $u$, we 
can substitute $u$ by $h$ to create an abstracted tactic.
We note this intermediate tactic $T[h/u]$. By repeating the process, multiple 
arguments may be abstracted in the same tactic. Ultimately, the more 
abstractions are performed the more general a tactic is, as many tactics 
become instances of the same abstracted tactic. Yet, it becomes harder and 
harder to predict suitable arguments.

\paragraph{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $A$, we first need to 
instantiate the placeholders $h$ inside $A$. Because it is difficult to 
build a general predictor that effective on each type of argument, we manually 
design different argument predictors for each type. Those predictors are 
usually given a goal has input and returns the best chosen argument for this 
goal. All in all, an abstracted tactics has 
dynamical arguments, it instantiate them differently for different goals.

@@

\paragraph{Selection of Abstracted Tactics}
As abstracted tactics do not appear in human proof, we need to find a way to 
predict them so that they can contribute to the \tactictoe proof search.
A straightforward idea is to try $A$ before $T$ during the proof search. 
However, this risks doing unnecessary work as they may have a similar effect. 
Therefore, we would like to decide 
beforehand if one is better than the other.
In fact, we can re-use the orthogonalization module to do this task for us.
We add $A$ to the competition, initially giving it the coverage of $T$. If $A$ 
wins, it becomes a a label for $g$ is including in the database of tactic 
features and thus can be predicted during proof search.
After many competitions, the coverage of $A$ may exceed the coverage of $T$. At 
this point, the coverage of $A$ is estimated on its own and not inherited from 
$T$ anymore.

@@

\paragraph{Reconstruction}
When a proof search succeeds (there are no more pending goals at the root)
we need to reconstruct a \holfour human-style proof.
The saved nodes consist of a set of trees where each edge is a tactic and
the proof tree is the one starting at the root.
In order to obtain a single \holfour proof, we need to combine the tactics
gathered in the trees using tacticals.
By the design of the search, a single tactic combinator, \texttt{THENL}, is 
sufficient. It combines a tactic with a list of subsequent ones, in such a way 
that after the parent tactic is called, for each created goal a respective 
tactic from the list is called.
The proof tree is transformed into a final single proof script
 by the following recursive function $P$ taking a
tree node $t$ and returning a string:
\begin{equation*}
P(t) =
\begin{cases}
P(c) & \text{if $t$ is a root},\\
tac & \text{if $t$ is a leaf},\\
tac\ \texttt{THENL}\ [P(c_0),\ldots,P(c_n)] & \text{otherwise.}
\end{cases}
\end{equation*}
where $tac$ is the tactic that produced the node, $c$ is the
only successful child of the root and $c_0, \ldots, c_n$ are the 
children of the node produced by the successful tactic.

>>>>

\paragraph{Abstraction of Tactic Arguments}
The first step is to abstract arguments of tactics by replacing them by a 
place holder creating an tactic with an unspecified argument.
Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can 
try to abstract any of the 
\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a placeholder $u$, we 
can substitute $u$ by $h$ to create an abstracted tactic.
We note this intermediate tactic $t[h/u]$. By repeating the 
process, multiple 
arguments may be abstracted in the same tactic. Ultimately, the more 
abstractions are performed the more general a tactic is, as many tactics 
become instances of the same abstracted tactic. Yet, it becomes harder and 
harder to predict suitable arguments.

\paragraph{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need 
to 
instantiate the placeholders $h$ inside $\hat{t}$. Because it is difficult to 
build a general predictor that effective on each type of argument, we manually 
design different argument predictors for each type. Those predictors are 
usually given a goal has input and returns the best chosen argument for this 
goal. All in all, an abstracted tactics has 
dynamical arguments, it instantiate them differently for different goals.

@@

\paragraph{Selection of Abstracted Tactics}
As abstracted tactics do not appear in human proof, we need to find a way to 
predict them so that they can contribute to the \tactictoe proof search.
A straightforward idea is to try $\hat{t}$ before $t$ during the proof search. 
However, this risks doing unnecessary work as they may have a similar effect. 
Therefore, we would like to decide 
beforehand if one is better than the other.
In fact, we can re-use the orthogonalization module to do this task for us.
We add $\hat{t}$ to the competition, initially giving it the coverage of $t$. 
If $\hat{t}$ 
wins, it becomes a a label for $g$ is including in the database of tactic 
features and thus can be predicted during proof search.
After many competitions, the coverage of $\hat{t}$ may exceed the coverage of 
$t$. At 
this point, the coverage of $\hat{t}$ is estimated on its own and not inherited 
from 
$t$ anymore.

@@

When \tactictoe finds a proof of a goal, it returns a search tree 
$\mathfrak{T}$ where the root node containing this goal is solved.
In order to transform this search tree into a proof script, we need to extract 
the tactics that contributed to the proof and combine them using tactic 
combinators.
\todoi{Define tactic combinators?}
By the design of the search, a single tactic combinator, \texttt{THENL}, is 
sufficient. It combines a tactic $t$ with a list of subsequent ones, in such a 
way that after $t$ is called, for each created goal a respective 
tactic from the list is called.
Here is a formal description of the transformation $Proof$ defined by mutual 
recursion on the goals and nodes of $\mathfrak{T}$:

%\begin{equation*}
%Proof(g) =_{def}
%\begin{cases}
%T(g) & \text{if $t$ is a root},\\
%tac & \text{if $t$ is a leaf},\\
%tac\ \texttt{THENL}\ [P(a_1),\ldots,t_{n-1}] & \text{otherwise.}
%\end{cases}
%
%
%\begin{cases}
%tac & \text{if $t$ is a root and $g$ the unique goal of the root},\\
%tac & \text{if $t$ is a leaf},\\
%tac\ \texttt{THENL}\ [P(a_1),\ldots,t_{n-1}] & \text{otherwise.}
%\end{cases}
%
%
%
%
%\end{equation*}
where $tac$ is the tactic that produced the node, $c$ is the
only successful child of the root and $c_0, \ldots, c_n$ are the 
children of the node produced by the successful tactic.
